---
title: "Reproducible_Document"
author: "MaibrittTornbjerg"
format: html
---

```{r setup}
#| message: false
#| warning: false


library(tidyverse)

post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)

tidied_post_meal_data <- tidier_post_meal_data |>
  mutate(
    insulin_glucose_ration = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First.degree relative")
  )
```

ved at ændre i vores setup chunk, sikrer vi os at de ændringer vi laver længere nede, bliver gemt og findes i vores enviorment, når vi skal bruge dem senere hen. 

## Writing R code 

2+2 

```{r}
2 + 2
```
## About me

-   Maibritt Egeskov Tornbjerg
-   Odense University Hospital

I work as a **reserch assistent** at the *KMEB* lab 
___

## Simple code”

```{r}
3 * 3
```
## Testing for Git

The wheather is sunny but windy

## Importing data 
 
importing data file "post-meal-insulin.csv" from a web page using the folowing line of code -> download.file(
  url = "https://zenodo.org/records/4989220/files/Eliasson_data2.csv?download=1", destfile = here::here("data/post-meal-insulin.csv"))

## Showing the data 
```{r}
post_meal_data

glimpse(post_meal_data)
```

## Plot one continuous variable
"na.rm = F" kan bruges til at fjerne manglende data 

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
see @fig-bmi-histo

## Exercise: discrete plots

```{r}
#| fig-cap: "Distribution of Glykemi"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

See @fig-glykemi-bar above for a cool plot!

## Plotting two screate variables 

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots

set up two plots side by side, when I render the document. 

```{r fig-bmi-glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|  - "Distribution of BMI"
#|  - "Number of those with glycemia"
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continous variables 

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg

ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Selecting data

Bruger "select" funktionen til at sortere data ud fra forskellige kreterier

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```

## Renaming Colums
```{r}
small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("..."))

# Task 3.
post_meal_data |>
  rename(
    P.Glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 4.
post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)
```

## Filtering data

```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")

small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```

## Modifying data

```{r}
tidier_post_meal_data |>
  mutate(insulin_glucose_ration = auc_insulin / auc_glucose)

# omdøbe variable, ved to forskellige variable. Her dem der tidligere hed CTR og FDR
tidier_post_meal_data |>
  mutate(
    insulin_glucose_ration = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First.degree relative")
  )
```

## Practice with filter and mutate

```{r}
# Task 1.
post_meal_data |>
  filter(BMI >= 25 & BMI <= 40 & Group == "FDR")

# Task 2.
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) |>
  filter(
    percent_body_fat > 30
  )
```


## Visualizing with a jitter plot
```{r}
tidied_post_meal_data |>
  mutate(Overweight = if_else(BMI >= 25, "Overweight", "Not Overweight")) |>
  ggplot(aes(x = Overweight, y = insulin_glucose_ration)) +
  geom_jitter()
```

## Plotting three or more varibles

kan ændre på faktorerne. De er default sat op alfabetisk, men dette kan ændres 
```{r}
tidied_post_meal_data |>
  filter(BMI < 30) |>
  mutate(young = if_else(Age < 40, "Young", "Old")) |>
  ggplot(aes(x = BMI, y = insulin_glucose_ration, colour = Group)) +
  geom_point() +
  facet_grid(rows = vars(young)) +
  labs(
    x = "Body mass index (BMI)",
    y = "Insulin to glucose ration"
  )
```

## Make multi-dimensional plots!

```{r}
#| label: fig-multidimensional
#| fig-cap: "Two, mult-dimensional plots of our data."
#| fig-subcap:
#|   - "Distribution of percent body fat by group and age category."
#|   - "Relationship between BMI and bone mineral density, split by insulin sensitivity."
#| layout-ncol: 2
# Task 1.
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100,
    young = if_else(Age < 40, "Old", "Young")
  ) |>
  ggplot(aes(x = Group, y = percent_body_fat, colour = young)) +
  geom_jitter(width = 0.2)

# Task 2.
post_meal_data |>
  mutate(
    insulin_glucose_ratio = (auc_ins / auc_pg),
    insulin_insensitive = if_else(insulin_glucose_ratio > mean(insulin_glucose_ratio), "Insensitive", "Sensitive")
  ) |>
  ggplot(aes(x = BMI, y = Bone.mineral.DXA)) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(insulin_insensitive))
```
